Missing The Point:

### 2020-07-15 ICML/MissingThePoint.Rmd
Crop shinymice hex image for presentation

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "ICML ARTEMISS workshop 2020"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Algorithmic Convergence

@@

```{r non-conv, echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
library(mice)
#load("example_diagnostics.Rdata")
load("../traceplotdata.Rdata")

# define colorblind friendly colors
paint5 <- c('#228833', '#66CCEE', '#CCBB44','#EE6677', '#AA3377')
# paint2 <- c('#66CCEE', '#EE6677')
# 
# # plot chain
# trace <- ggplot(diagnostics) +
#   geom_point(aes(x = iteration, y = `Chain 1`, color = patho),
#              size = .5,
#              na.rm = TRUE) +
#   geom_point(aes(x = iteration, y = `Chain 2`, color = patho),
#              size = .5,
#              na.rm = TRUE) +
#   geom_point(aes(x = iteration, y = `Chain 3`, color = patho),
#              size = .5,
#              na.rm = TRUE) +
#   geom_point(aes(x = iteration, y = `Chain 4`, color = patho),
#              size = .5,
#              na.rm = TRUE) +
#   geom_point(aes(x = iteration, y = `Chain 5`, color = patho),
#              size = .5,
#              na.rm = TRUE) +
#   geom_line(aes(x = iteration, y = `Chain 2`, color = patho), size = .5, na.rm = TRUE) +
#   geom_line(aes(x = iteration, y = `Chain 3`, color = patho), size = .5, na.rm = TRUE) +
#   geom_line(aes(x = iteration, y = `Chain 1`, color = patho), size = .5, na.rm = TRUE) +
#   geom_line(aes(x = iteration, y = `Chain 4`, color = patho), size = .5, na.rm = TRUE) +
#   geom_line(aes(x = iteration, y = `Chain 5`, color = patho), size = .5, na.rm = TRUE) +
#   scale_colour_manual(values=paint2) +
#   scale_x_continuous(breaks = 1:10) +
#   xlab("Iteration") +
#   ylab("Chain mean") + #bquote(theta)) +
#   labs(color = "") + 
#   theme_classic() +
#   theme(legend.position = "bottom")

#ggplotly(trace)

# alternatively, use the actual data
# run 1.Execute.R until imputing anything, than run RunSimulationOnce.R until after amputing the data
# with the 'amps' object
set.seed(12)
d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
#plot(d75)
p <- gg.mids(d75, x="Y")[[1]] 
p <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    x = "Iteration",
    y = "Chain mean", 
    colour = "Imp.")

ggplotly(p)

```
---

# Simulation Study

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=1000$, $n_{\rm sim} = 1000$), we obtain valid regression estimates after 5 to 10 iterations. <!-- Multivariate normal set ( $n = 1000$ ), 'amputed' with a varying proportion of incomplete cases, and 'imputed' with a varying number of iterations. Depicted are the bias and coverage rate of an estimated regression coefficient ( $n_{\rm sim} = 1000$ ). -->

@@

est_Rh <- results %>% ggplot(aes(x = t, y = r.hat.max.beta, color = as.factor(p*100))) +
  geom_hline(yintercept = 1,
             color = "grey",
             lwd = 
               1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.2), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.1), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.01), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  scale_y_continuous(limits = c(1,1.53), breaks = c(1,1.1,1.2,1.3,1.4,1.5)) +
  labs(x = "Number of iterations", 
       y = "Ř", #r'$\\widehat{R}$'
       colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

@@

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_cov)
```
]

---

# Simulation Study (2)

Whereas non-convergence diagnostics $\widehat{R}$ and auto-correlation identify signs of non-convergence up-to 30 to 50 iterations.

@@

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_AC)
```
]

---
# Take-away

@@

.pull-left[
<br> <br> <br>
<br> <br> <br>
<br> <br> <br>
Upcoming...
]

.pull-right[
<img src="https://raw.githubusercontent.com/gerkovink/shinyMice/2dae400d9d23ede7bec031c1501acdc681eab13c/ThesisProposal/Figures/logo.png" width="420">
]

<!-- .footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).] -->

>>>>

---
title: "Missing The Point:"
subtitle: "Non-Convergence in Iterative Imputation Algorithms"
author: "Hanne I. Oberman, Stef van Buuren, Gerko Vink"
institute: "ICML ARTEMISS workshop 2020"
date: "17-07-2020" #(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    #seal: false
    self_contained: true
    css: [default, default-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      ratio: "16:9"
---

```{css echo=FALSE}
.right-column{
  padding-top: 0;
}
```

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)

# set-up env
library(dplyr)
library(ggplot2)
library(plotly)
library(mice)
#load("example_diagnostics.Rdata")
# load("../traceplotdata.Rdata")
# set.seed(12)
# d75 <- amps[[4]] %>% mice(maxit=10, printFlag = FALSE)
# save(d75, file = "imps.Rdata")
load("imps.Rdata")

# define colorblind friendly colors
paint5 <- c('#228833', '#66CCEE', '#CCBB44','#EE6677', '#AA3377')
# paint2 <- c('#66CCEE', '#EE6677')
```

.left-column[ 
```{r echo=TRUE, eval=FALSE}

Missing The Point

Non-Convergence 
In Iterative 
Imputation 
Algorithms

Hanne I. Oberman, 
Stef van Buuren, 
& Gerko Vink

ARTEMISS workshop
ICML 2020

]

.right-column[ 
# Algorithmic Convergence

@@

```{r non-conv, echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
p <- gg.mids(d75, x="Y")[[1]] 
p_up <- p + 
  scale_color_manual(values = paint5) + 
  scale_x_continuous(breaks = 1:10) + 
  labs(
    #title = "Traceplot", 
    x = "Iteration",
    y = "Chain mean", 
    colour = "Imp.")

```{r echo=FALSE, fig.width=10, fig.height = 5.2, message=FALSE, warning=FALSE}
ggplotly(p_up)
```
]
---

.left-column[ 
```{r echo=TRUE, eval=FALSE}

Missing The Point

Non-Convergence 
In Iterative 
Imputation 
Algorithms

Hanne I. Oberman, 
Stef van Buuren, 
& Gerko Vink

ARTEMISS workshop
ICML 2020

]

.right-column[ 
# Simulation Study

When imputing an incomplete multivariate normal set ( $n_{\rm obs}=$ 1000, $n_{\rm sim}=$ 1000), we obtain valid regression estimates after 5 to 10 iterations. <!-- Multivariate normal set ( $n = 1000$ ), 'amputed' with a varying proportion of incomplete cases, and 'imputed' with a varying number of iterations. Depicted are the bias and coverage rate of an estimated regression coefficient ( $n_{\rm sim} = 1000$ ). -->

@@

est_Rh <- results %>% ggplot(aes(x = t, y = r.hat.max.beta, color = as.factor(p*100))) +
  geom_hline(yintercept = 1,
             color = "grey",
             lwd = 
               1) +
  geom_point(size = .5, na.rm = TRUE) +
  geom_line(size = .5, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.2), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.1), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  geom_line(aes(x=t, y=thresh1.01), color = "grey", linetype = "dashed", size = .25, na.rm = TRUE) +
  scale_colour_manual(values=paint5) +
  scale_y_continuous(limits = c(1,1.53), breaks = c(1,1.1,1.2,1.3,1.4,1.5)) +
  labs(x = "Number of iterations", 
       y = "Potential scale reduction factor", #r'$\\widehat{R}$'
       colour = "Miss. (%)") +
  theme_classic() +
  theme(legend.position = "bottom")

@@

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_cov)
```
]
]
---

.left-column[ 
```{r echo=TRUE, eval=FALSE}

Missing The Point

Non-Convergence 
In Iterative 
Imputation 
Algorithms

Hanne I. Oberman, 
Stef van Buuren, 
& Gerko Vink

ARTEMISS workshop
ICML 2020

]

.right-column[ 
# Simulation Study (2)

Whereas non-convergence diagnostics (potential scale reduction factor, auto-correlation) identify signs of non-convergence up-to 30 to 50 iterations.

@@

.pull-right[
```{r echo=FALSE, fig.width=5.1, fig.height=5.4, message=FALSE, warning=FALSE}
ggplotly(est_AC)
```
]
]
---

.left-column[ 
```{r echo=TRUE, eval=FALSE}

Missing The Point

Non-Convergence 
In Iterative 
Imputation 
Algorithms

Hanne I. Oberman, 
Stef van Buuren, 
& Gerko Vink

ARTEMISS workshop
ICML 2020

]

.right-column[ 
# Take-away

@@

.pull-left[
<br> <br> <br>
<br> <br> <br>
<br> <br> <br>
<br> <br> <br>
Upcoming...
]

.pull-right[
<img src="https://raw.githubusercontent.com/gerkovink/shinyMice/2dae400d9d23ede7bec031c1501acdc681eab13c/ThesisProposal/Figures/logo.png" width="595">
]
]
<!-- .footnote[Slides created via the R package [**xaringan**](https://github.com/yihui/xaringan).] -->
